\morus: Biases in the Stream

### 2018-05-09 morusAC_04_Mini_trails.tex
propose introduction (poorly written)
Describe attack on MORUS1280 primarly

\section{Linear Trail for \cipher{MiniMORUS}}

@@

\begin{figure}
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{subfigure}{.32\textwidth}
  \centering
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, alpha]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
      ;
    \draw[trail, alpha, dashed]
      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
      ;
  \end{tikzpicture}%}}}
  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \centering
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, beta]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (tlll-1) -- (W-20) node[above] {$i$}
      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
      (xnd-1) -- (and-1)
      (W40) node[below] {\phantom{$i$}}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, gamma]
      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
      (xnd1) -- (and1)
      (xor1) -- (txor1)
      (txor1) -- (W-24) node[above] {$i$}
      (lll1) -- (W41) node[below] {$i+b_1$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
  \end{subfigure}
  \bigskip

  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, delta]
      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
      (lll4) -- (W44) node[below] {$i+b_4$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.75,yscale=1.5]%{{{
    \printstate
    \draw[trail, epsil]
      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
      (lll2) -- (W42) node[below] {$i+b_2$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
    \small
    \begin{align*}
      \intertext{\cipher{MiniMORUS-640}:}
\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
      \intertext{\cipher{MiniMORUS-1280}:}
\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
    \end{align*}
    \caption*{\cipher{MiniMORUS} instances}
  \end{subfigure}
  \caption{\cipher{MiniMORUS} linear trail fragments.}
  \label{fig:minimorusfragments}
\end{figure}

@@

In order to build the equation for the full trail, we start with $E^2_0$:
\[
S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
\]
In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$). This yields:
\begin{align*}
&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
\end{align*}
We then need to cancel two terms of the form $S^t_{2,i}$. To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$. This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms. By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms. By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms). Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits. Ultimately, for \cipher{MiniMORUS640}, this yields the equation:

\begin{flalign*}
&&&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}&&\\
&&\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}&&\\
&&=\; &S^0_{1,27} \oplus S^0_{3,27} \oplus S^0_{1,27} \cdot S^0_{2,27} \oplus S^0_{2,27} \cdot S^0_{3,27}&&\text{weight 1}\\
&&\oplus\; &S^1_{1,26} \oplus S^1_{3,26} \oplus S^1_{1,26} \cdot S^1_{2,26} \oplus S^1_{2,26} \cdot S^1_{3,26}&&\text{weight 1}\\
&&\oplus\; &S^1_{1,8} \oplus S^1_{3,8} \oplus S^1_{1,8} \cdot S^1_{2,8} \oplus S^1_{2,8} \cdot S^1_{3,8}&&\text{weight 1}\\
&&\oplus\; &S^2_{3,1} \oplus S^2_{1,1} \oplus S^2_{1,1} \cdot S^2_{2,1} \oplus S^2_{2,1} \cdot S^2_{3,1}&&\text{weight 1}\\
&&\oplus\; &S^2_{1,7} \oplus S^2_{3,7} \oplus S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7}&&\text{weight 1}\\
&&\oplus\; &S^2_{1,15} \oplus S^2_{3,15} \oplus S^2_{1,15} \cdot S^2_{2,15} \oplus S^2_{2,15} \cdot S^2_{3,15}&&\text{weight 1}\\
&&\oplus\; &S^2_{1,27} \oplus S^2_{3,27} \oplus S^2_{1,27} \cdot S^2_{2,27} \oplus S^2_{2,27} \cdot S^2_{3,27}&&\text{weight 1}\\
&&\oplus\; &S^1_{1,2} \oplus S^1_{3,2} \oplus S^1_{1,2} \cdot S^1_{2,2} \oplus S^1_{2,2} \cdot S^1_{3,2}&&\text{weight 1}\\
&&\oplus\; &S^3_{3,14} \oplus S^3_{1,14} \oplus S^3_{1,14} \cdot S^3_{2,14} \oplus S^3_{2,14} \cdot S^3_{3,14}&&\text{weight 1}\\
&&\oplus\; &S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
&&\oplus\; &S^2_{2,31} \cdot S^2_{3,31}&&\text{weight 1}\\
&&\oplus\; &S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
&&\oplus\; &S^3_{0,7} \cdot S^3_{1,7}&&\text{weight 1}\\
&&\oplus\; &S^3_{2,6} \cdot S^3_{3,6}&&\text{weight 1}\\
&&\oplus\; &S^3_{2,12} \cdot S^3_{3,12}&&\text{weight 1}\\
&&\oplus\; &S^4_{2,19} \cdot S^4_{3,19}.&&\text{weight 1}
\end{flalign*}

The equation for \cipher{MiniMORUS1280} is very similar, and is given in \Cref{sec:traileq}.

@@

In the equation for \cipher{MiniMORUS640} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1. By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16. Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$. The same holds for \cipher{MiniMORUS1280} (cf. \Cref{sec:traileq}).

@@

To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}. In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$). Results are displayed on \Cref{tab:miniapproximations}.
While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}. The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.

@@

>>>>

\section{Linear Trail for \cipher{MiniMORUS}}
\label{sec/minitrails}

In this section, we describe how we build a trail for \cipher{MiniMORUS} before computing its bias and experimentally verify it.

@@

\begin{figure}
  \substatesfalse
  % \substatesfalse to label state words and/or masks
  \centering
  \begin{subfigure}{.32\textwidth}
  \centering
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, alpha]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (xnd-1) -- (and-1) (xor-1) -- (txor-1) (txor-1) -- (tanB0) (tanB0) -- (tanB0-|and0.east)
      (and0) -- (xnd0) (xnd0) -- (tlll-1) (xnd0) -- (xor0)
      (and-1.east|-tanA-1) -- (tanA-1) (tanA-1) -- (txor0) (txor0) -- (xor0)
      (xor0) -- (lll0) (lll0) -- (W40) node[below] {$i+b_0$}
      ;
    \draw[trail, alpha, dashed]
      (and-1.east|-tanB-1) -- (tanB-1) (tanB-1) -- (tanA0) (tanA0) -- (tanA0-|and0.east)
      ;
  \end{tikzpicture}%}}}
  \caption*{$\alpha^t_i$: weight 1 (not 2)} %$C^j_i, S^{j-1,0}_{i+5}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \centering
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, beta]
      (C) -- node[right] {$i$} (lll-1)
      (lll-1) -- (tlll-1) (lll-1) -- (xor-1) (xor-1) -- (xnd-1)
      (tlll-1) -- (W-20) node[above] {$i$}
      (xor-1) -- (txor-1) (txor-1) -- (W-21) node[above] {$i$}
      (xnd-1) -- (and-1)
      (W40) node[below] {\phantom{$i$}}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\beta^t_i$: weight 1} %$C^j_i, S^{j,0}_i, S^{j,1}_i$ ($w\!=\!0$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, gamma]
      (W-21) node[above] {$i$} -- (xnd1) (xnd1) -- (xor1) (xor1) -- (lll1)
      (xnd1) -- (and1)
      (xor1) -- (txor1)
      (txor1) -- (W-24) node[above] {$i$}
      (lll1) -- (W41) node[below] {$i+b_1$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\gamma^t_i$: weight 1} %$S^{j,1}_i, S^{j,4}_i, S^{j+1,1}_{i-1}$ ($w\!=\!1$)}
  \end{subfigure}
  \bigskip

  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, delta]
      (W-24) node[above] {$i$} -- (xnd4) (xnd4) -- (xor4) (xnd4) -- (and4)
      (xor4) -- (lll4) (xor4) -- (txor4) (txor4) -- (W42) node[below] {$i$}
      (lll4) -- (W44) node[below] {$i+b_4$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\delta^t_i$: weight 1} %$S^{j,4_i}, S^{j+1,2}_i, S^{j+1,4}_{i+13}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
  \begin{tikzpicture}[xscale=0.65,yscale=1.5]%{{{
    \printstate
    \draw[trail, epsil]
      (W-22) node[above] {$i$} -- (xnd2) (xnd2) -- (xor2) (xnd2) -- (and2)
      (xor2) -- (lll2) (xor2) -- (txor2) (txor2) -- (W40) node[below] {$i$}
      (lll2) -- (W42) node[below] {$i+b_2$}
      ;
  \end{tikzpicture}%}}}
  \caption*{$\varepsilon^t_i$: weight 1} %$S^{j,2}_i, S^{j+1,0}_i, S^{j+1,2}_{i+7}$ ($w\!=\!1$)}
  \end{subfigure}
  \hfill
  \begin{subfigure}{.32\textwidth}
    \small
    \begin{align*}
      \intertext{\cipher{MiniMORUS-640}:}
\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+5} \\
\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+31} &\to S^t_{4,i} \\
\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+13} &\to S^{t+1}_{2,i} \\
\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+7} &\to S^{t+1}_{0,i}.
      \intertext{\cipher{MiniMORUS-1280}:}
\alpha^t_i:&& C^t_i &\to S^{t+1}_{0,i+13} \\
\beta^t_i:&& C^t_i, S^t_{0,i} &\to S^t_{1,i} \\
\gamma^t_i:&& S^t_{1,i}, S^{t+1}_{1,i+46} &\to S^t_{4,i} \\
\delta^t_i:&& S^t_{4,i}, S^{t+1}_{4,i+4} &\to S^{t+1}_{2,i} \\
\varepsilon^t_i:&& S^t_{2,i}, S^{t+1}_{2,i+38} &\to S^{t+1}_{0,i}.
    \end{align*}
    \caption*{\cipher{MiniMORUS} instances}
  \end{subfigure}
  \caption{\cipher{MiniMORUS} linear trail fragments.}
  \label{fig:minimorusfragments}
\end{figure}

@@

In order to build the equation for the full trail, we start with $E^2_0$:
\[
S^2_{2,0} \oplus S^{3}_{2,b_2} \oplus S^{3}_{0,0} = S^2_{3,0} \cdot S^2_{4,0}.
\]
In order to cancel the $S^{3}_{0,0}$ term on the left-hand side, we add in the equation $A^2_{-b_0}$ (where the sum of two equations of the form $a = b$ and $c = d$ is defined to be $a+c = b+d$). This yields:
\begin{align*}
&S^2_{2,0} \oplus S^3_{2,b_2} \oplus C^2_{-b_0}\\
=\; &S^2_{3,0} \cdot S^2_{4,0} \oplus S^2_{1,-b_0} \oplus S^2_{3,-b_0} \oplus S^2_{1,-b_0} \cdot S^2_{2,-b_0} \oplus S^2_{2,-b_0} \cdot S^2_{3,-b_0}.
\end{align*}
We then need to cancel two terms of the form $S^t_{2,i}$. To do this, we add in the equations $D^t_i$ for appropriate choices of $t$ and $i$. This replaces the two $S^t_{2,i}$ terms by four $S^t_{4,i}$ terms. By using equation $B^t_i$ four times, we can then replace these four $S^t_{4,i}$ terms by eight $S^t_{1,i}$ terms. By applying equation $B^t_i$ eight times, these eight $S^t_{1,i}$ terms can in turn be replaced by eight $S^t_{0,i}$ terms (and some ciphertext terms). Finally, applying $A^t_i$ eight times allows us to replace these eight $S^t_{0,i}$ terms by only ciphertext bits. Ultimately, for \cipher{MiniMORUS1280}, this yields the equation:
\begin{flalign*}
&&&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}&&\\
&&\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}&&\\
&&=\; &S^0_{1,51} \cdot S^0_{2,51} \oplus S^0_{2,51} \cdot S^0_{3,51} \oplus S^0_{1,51} \oplus S^0_{3,51}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,25} \cdot S^1_{2,25} \oplus S^1_{2,25} \cdot S^1_{3,25} \oplus S^1_{1,25} \oplus S^1_{3,25}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,33} \cdot S^1_{2,33} \oplus S^1_{2,33} \cdot S^1_{3,33} \oplus S^1_{1,33} \oplus S^1_{3,33}&&\text{weight 1}\\
&&\oplus\; & S^1_{1,55} \cdot S^1_{2,55} \oplus S^1_{2,55} \cdot S^1_{3,55} \oplus S^1_{1,55} \oplus S^1_{3,55}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,7} \cdot S^2_{2,7} \oplus S^2_{2,7} \cdot S^2_{3,7} \oplus S^2_{1,7} \oplus S^2_{3,7}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,29} \cdot S^2_{2,29} \oplus S^2_{2,29} \cdot S^2_{3,29} \oplus S^2_{1,29} \oplus S^2_{3,29}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,37} \cdot S^2_{2,37} \oplus S^2_{2,37} \cdot S^2_{3,37} \oplus S^2_{1,37} \oplus S^2_{3,37}&&\text{weight 1}\\
&&\oplus\; & S^2_{1,51} \cdot S^2_{2,51} \oplus S^2_{2,51} \cdot S^2_{3,51} \oplus S^2_{1,51} \oplus S^2_{3,51}&&\text{weight 1}\\
&&\oplus\; & S^3_{1,11} \cdot S^3_{2,11} \oplus S^3_{2,11} \cdot S^3_{3,11} \oplus S^3_{1,11} \oplus S^3_{3,11}&&\text{weight 1}\\
&&\oplus\; & S^2_{0,0} \cdot S^2_{1,0}&&\text{weight 1}\\
&&\oplus\; & S^2_{2,46} \cdot S^2_{3,46}&&\text{weight 1}\\
&&\oplus\; & S^2_{3,0} \cdot S^2_{4,0}&&\text{weight 1}\\
&&\oplus\; & S^3_{0,38} \cdot S^3_{1,38}&&\text{weight 1}\\
&&\oplus\; & S^3_{2,20} \cdot S^3_{3,20}&&\text{weight 1}\\
&&\oplus\; & S^3_{2,50} \cdot S^3_{3,50}&&\text{weight 1}\\
&&\oplus\; & S^4_{2,24} \cdot S^4_{3,24}&&\text{weight 1}
\end{flalign*}

The equation for \cipher{MiniMORUS640} is very similar, and is given in \Cref{sec:traileq}.

@@

In the equation for \cipher{MiniMORUS1280} from the previous section, each line on the right-hand side of the equality involves distinct $S^t_{i,j}$ terms, and each line has a weight of 1. By the piling-up lemma, it follows that if we assume distinct $S^t_{i,j}$ terms to be uniform and independent, then the expression on the right-hand side has a weight of 16. Hence the linear combination of ciphertext bits on the left-hand side has an expected bias of $2^{-16}$. The same holds for \cipher{MiniMORUS640} (cf. \Cref{sec:traileq}).

@@

To confirm that our analysis is correct, we ran experiments on an implementation of \cipher{MiniMORUS1280} and \cipher{MiniMORUS640}. In both cases, we measured the bias of two halves $\chi_1$ and $\chi_2$ of the full trail, as well as the bias of the full trail itself (denoted by $\chi$). Results are displayed on \Cref{tab:miniapproximations}.
While our analysis from the previous section predicts a bias of $2^{-16}$, experiments indicate a slightly higher bias of $2^{-15.5}$ for \cipher{MORUS640}. The discrepancy of $2^{-0.5}$ probably arises from the fact that register bits across different steps are not completely independent.

@@

%%% Local Variables:
%%% TeX-master: "morusAC"
%%% End:

### 2018-05-09 morusAC_05_Full_trails.tex
Describe attack on MORUS1280 primarly

In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS640} and \cipher{MiniMORUS1280}. We now turn to the full ciphers \cipher{MORUS640} and \cipher{MORUS1280}.

\subsection{Symmetrizing the Trail}
In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
That is, for every $S^t_{i,j}$ (resp. $C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp. $C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size. For example, if $w=32$ (for \cipher{MORUS640}), the term $S^3_{2,0}$ is symmetrized into:
\[
S^3_{2,0} \oplus S^3_{2,32} \oplus S^3_{2,64} \oplus S^3_{2,96}.
\]

Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
In this way, the biased linear combination on \cipher{MiniMORUS640} given in \Cref{sec:minitraileq}, namely:
\begin{align*}
&C^0_{27} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^2_{1} \oplus C^2_{13} \oplus C^2_{15}\\
\oplus\; &C^2_{27} \oplus C^2_{31} \oplus C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^4_{19}
\end{align*}
ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS640}:
\begin{align*}
&C^0_{27} \oplus C^0_{59} \oplus C^0_{91} \oplus C^0_{123} \oplus C^1_{0} \oplus C^1_{2} \oplus C^1_{8} \oplus C^1_{26} \oplus C^1_{32} \oplus C^1_{34}\\
\oplus\; & C^1_{40} \oplus C^1_{58} \oplus C^1_{64} \oplus C^1_{66} \oplus C^1_{72} \oplus C^1_{90} \oplus C^1_{96} \oplus C^1_{98} \oplus C^1_{104} \oplus C^1_{122}\\
\oplus\; & C^2_{1} \oplus C^2_{13} \oplus C^2_{15} \oplus C^2_{27} \oplus C^2_{31} \oplus C^2_{33} \oplus C^2_{45} \oplus C^2_{47} \oplus C^2_{59} \oplus C^2_{63}\\
\oplus\; & C^2_{65} \oplus C^2_{77} \oplus C^2_{79} \oplus C^2_{91} \oplus C^2_{95} \oplus C^2_{97} \oplus C^2_{109} \oplus C^2_{111} \oplus C^2_{123} \oplus C^2_{127}\\
\oplus\; & C^3_{6} \oplus C^3_{12} \oplus C^3_{14} \oplus C^3_{20} \oplus C^3_{38} \oplus C^3_{44} \oplus C^3_{46} \oplus C^3_{52} \oplus C^3_{70} \oplus C^3_{76}\\
\oplus\; & C^3_{78} \oplus C^3_{84} \oplus C^3_{102} \oplus C^3_{108} \oplus C^3_{110} \oplus C^3_{116} \oplus C^4_{19} \oplus C^4_{51} \oplus C^4_{83} \oplus C^4_{115}.
\end{align*}
We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS1280}.

@@

Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS640} and \cipher{MORUS1280}. However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.

>>>>

In the previous section, we have exhibited a linear trail on the reduced ciphers \cipher{MiniMORUS1280} and \cipher{MiniMORUS640}. We now turn to the full ciphers \cipher{MORUS1280} and \cipher{MORUS640}.

\subsection{Symmetrizing the Trail}
In order to build a trail on the full \cipher{MORUS}, we proceed exactly like we did for \cipher{MiniMORUS}, following the exact same path down to step and word rotation values, with one difference: in order to move from the one-word registers of \cipher{MiniMORUS} to the four-word registers of full \cipher{MORUS}, we make every term $S^t_{i,j}$ and $C^t_j$ symmetric, in the sense of \Cref{sec/introminimorus}.
That is, for every $S^t_{i,j}$ (resp. $C^t_j$) component in every trail fragment and every equation, we \emph{symmetrize} the term by adding in the terms $S^t_{i,j+w}$, $S^t_{i,j+2w}$, $S^t_{i,j+3w}$ (resp. $C^t_{j+w}$, $C^t_{j+2w}$, $C^t_{j+3w}$), where as usual $w$ denotes the word size. For example, if $w=64$ (for \cipher{MORUS1280}), the term $S^3_{2,0}$ is symmetrized into:
\[
S^3_{2,0} \oplus S^3_{2,64} \oplus S^3_{2,128} \oplus S^3_{2,192}.
\]

Thus, translating the trail from one of the \cipher{MiniMORUS} ciphers to the corresponding full \cipher{MORUS} cipher amounts to making every linear combination symmetric---indeed, that was the point of introducing \cipher{MiniMORUS} in the first place.
Concretely, in order to build the full trail equation for \cipher{MORUS}, we write symmetric versions of equations $A^t_i$, $B^t_i$, $C^t_i$, $D^t_i$, $E^t_i$ from \Cref{sec:minitraileq}, and then combine them in exactly the same way as before.
In this way, the biased linear combination on \cipher{MiniMORUS1280} given in \Cref{sec:minitraileq}, namely:
\begin{align*}
&C^0_{51} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37}\\
\oplus\; &C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^3_{11} \oplus C^3_{20} \oplus C^3_{42} \oplus C^3_{50} \oplus C^4_{24}
\end{align*}
ultimately yields the following biased symmetrized linear combination on the full \cipher{MORUS1280}:
\begin{align*}
&C^0_{51} \oplus C^0_{115} \oplus C^0_{179} \oplus C^0_{243} \oplus C^1_{0} \oplus C^1_{25} \oplus C^1_{33} \oplus C^1_{55} \oplus C^1_{64} \oplus C^1_{89}\\
\oplus\; & C^1_{97} \oplus C^1_{119} \oplus C^1_{128} \oplus C^1_{153} \oplus C^1_{161} \oplus C^1_{183} \oplus C^1_{192} \oplus C^1_{217} \oplus C^1_{225} \oplus C^1_{247}\\
\oplus\; & C^2_{4} \oplus C^2_{7} \oplus C^2_{29} \oplus C^2_{37} \oplus C^2_{38} \oplus C^2_{46} \oplus C^2_{51} \oplus C^2_{68} \oplus C^2_{71} \oplus C^2_{93}\\
\oplus\; & C^2_{101} \oplus C^2_{102} \oplus C^2_{110} \oplus C^2_{115} \oplus C^2_{132} \oplus C^2_{135} \oplus C^2_{157} \oplus C^2_{165} \oplus C^2_{166} \oplus C^2_{174}\\
\oplus\; & C^2_{179} \oplus C^2_{196} \oplus C^2_{199} \oplus C^2_{221} \oplus C^2_{229} \oplus C^2_{230} \oplus C^2_{238} \oplus C^2_{243} \oplus C^3_{11} \oplus C^3_{20}\\
\oplus\; & C^3_{42} \oplus C^3_{50} \oplus C^3_{75} \oplus C^3_{84} \oplus C^3_{106} \oplus C^3_{114} \oplus C^3_{139} \oplus C^3_{148} \oplus C^3_{170} \oplus C^3_{178}\\
\oplus\; & C^3_{203} \oplus C^3_{212} \oplus C^3_{234} \oplus C^3_{242} \oplus C^4_{24} \oplus C^4_{88} \oplus C^4_{152} \oplus C^4_{216}
\end{align*}
We refer the reader to \Cref{sec:traileq} for the corresponding linear combination on \cipher{MORUS640}.

@@

Following the previous intuition, one may expect that the weight of the symmetrized trail should simply be four times the weight of the corresponding \cipher{MiniMORUS} trail, namely 64 for both \cipher{MORUS1280} and \cipher{MORUS640}. However, reality is a little more complex, as the symmetrized trail does not exactly behave as four copies of the original trail when one considers nonlinear terms.
